Interactive Visualization of Hierarchically Structured Data

### 2016-12-04 treelapse.Rmd
echo = FALSE, and revisions for first few paragraphs

```{r set-chunk-opts}
knitr::opts_chunk$set(fig.align = "center", out.width = "200px")
```

@@

In this work, we describe methods for visualization of data structured along
trees, especially hierarchically structured collections of time
series. Influenced by Degree of Interest trees, linked scatterplot brushing,
and timeboxes, we have adapted the fundamental visualization principles of
"focus + context" and "linking" to study questions that arise when analyzing
tree structured data. Our motivating application is to the analysis of
microbial time series, where an evolutionary tree relating microbes is
available a priori. However, we have identified common problem types where, if
a tree is not directly available, it can be constructed from data and then
studied using our techniques.

We have implemented our visualizations in d3, but encapsulated in an R package,
to facilitate rapid turnover from data preparation and modeling to interactive
exploration, and vice versa. We hope this package encourages data analysts to
work at this border of data modeling and visualization, and more generally
empowers a wider audience to apply less widely known, but powerful,
visualization ideas.

>>>>

```{r set-chunk-opts}
knitr::opts_chunk$set(fig.align = "center", out.width = "200px", echo = FALSE)
```

@@

(todo: sentence about why we should care, and a high level summary
before diving into degree-of-interest linking jargon)

In this work, we describe methods for visualization of data structured
along trees, especially hierarchically structured collections of time
series. Influenced by Degree-of-Interest trees [@heer2004doitrees],
linked scatterplot brushing [@becker1987brushing], and timeboxes
[@hochheiser2004dynamic], we have adapted the fundamental
visualization principles of "focus + context" and "linking" to study
questions that arise when analyzing tree structured data. Our
motivating application is to the analysis of microbial time series,
where an evolutionary tree relating microbes is available a
priori. However, we have identified common problem types where, if a
tree is not directly available, it can be constructed from data and
then studied using our techniques.

We have implemented our visualizations in d3, but encapsulated in an R
package, to facilitate rapid turnover from data preparation and
modeling to interactive exploration, and vice versa. We hope this
package encourages data analysts to work at the border between data
modeling and visualization, and more generally empowers a wider
audience to apply less widely known, but powerful, visualization
ideas.
